The single-asset analysis uses daily SPY (SPDR S\&P~500 ETF) VWAP prices from January~3, 2014 to January~3, 2024, a ten-year in-sample (IS) window of $\TIS{2{,}516}$ trading days; January~4, 2024 through April~20, 2026 is held out for out-of-sample (OoS) evaluation, giving $\TOoS{572}$ growth rates for the generator diagnostics. The VaR back-tests additionally prepend the boundary growth rate into the first held-out session, so their forecast count is $573$. Prices are split-adjusted but not dividend-adjusted, from commercial daily-price vendors. Cross-ticker generalisation uses a sector-balanced 30-ticker panel spanning ten GICS sectors with three large-cap representatives each, fit one CHMM per ticker.

The CHMM is benchmarked against one representative per model class, all fit on the same IS window: a plain i.i.d.\ empirical bootstrap~\cite{efron1979bootstrap} as the strongest non-parametric baseline; GARCH(1,1) with Gaussian~\cite{bollerslev1986generalized} and Student-$t$~\cite{bollerslev1987conditionally} innovations; MS-GARCH~\cite{haas2004new} as the closest regime-switching comparator, with a Bayesian re-run via the \texttt{MSGARCH} package~\cite{ardia2019msgarch}; a maximum-likelihood HSMM-N as the explicit-duration reference; and an in-house QuantGAN~\cite{wiese2020quantgan} as the deep-generative negative control. For every model we simulate $1{,}000$ independent paths matching the scored window~\cite{stenger2024thinking}, under deterministic seeds.

We score three primary metrics. The two-sample Kolmogorov--Smirnov (KS) pass rate at $\alpha = 0.05$ counts the fraction of simulated paths whose marginal is not rejected against the observed window; because the asymptotic KS null assumes i.i.d.\ observations while both series carry serial dependence, we read this pass rate as a descriptive marginal-fidelity score rather than a calibrated test. Mean simulated excess kurtosis is read against the observed value on the same window. The mean absolute error between observed and simulated ACFs of absolute growth rates over $L = 252$ lags, computed per path and averaged,
\begin{equation}
    \text{ACF-MAE} = \frac{1}{P}\sum_{p=1}^{P}\frac{1}{L}\sum_{\tau=1}^{L} \left| \hat\rho^{\text{obs}}_{|G|}(\tau) - \hat\rho^{\,\text{sim},(p)}_{|G|}(\tau) \right|,
    \label{eq:acf_mae}
\end{equation}
captures volatility clustering, the third Cont stylized fact~\cite{cont2001empirical}. The VaR panel reports the unconditional Kupiec coverage test~\cite{kupiec1995techniques}, the Christoffersen joint conditional-coverage test~\cite{christoffersen1998evaluating} as the primary backtest, and the Engle--Manganelli dynamic-quantile (DQ) test~\cite{engle2004caviar}, against filtered-bootstrap~\cite{barone1999var} and CAViaR~\cite{engle2004caviar} contenders on the same harness. The cross-generator ranking is robust to the distributional-metric choice across KS, Anderson--Darling, Hellinger, and Wasserstein-1 panels; the continuous ranked probability score (CRPS)~\cite{diebold1995comparing} is reported as a tie-break.
